Robust co-occurrence quantification for low-dimensional lexical distributional semantics

### 2016-02-29 paper.tex
Matt

In contrast, much work on \emph{compositional} distributional semantics employs vectors with much fewer dimensions: e.g.~\textbf{2K} \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, \textbf{3K} \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or \textbf{10K} \cite{polajnar-clark:2014:EACL,Baroni2010nouns}. These models  assign tensors to functional words; for a vector space $V$ with dimension $k$,  a tensor $V \otimes V \cdots \otimes V$ of rank $n$ has dimension  $n^k$. As intransitive verbs and adjectives have  tensors of rank 2,  transitive verbs  rank 3, and prepositions rank 7, 
if one takes $n= \textbf{30k}$ one will end up with   spaces with intractable dimensions  $\textbf{30}^\textbf{2} \textbf{k}$, $\textbf{30}^\textbf{3} \textbf{k}$ and $\textbf{30}^\textbf{7} \textbf{k}$.

@@

Most co-occurrence weighting schemes in distributional semantics are based on \emph{point-wise mutual information} (PMI, Equation~\ref{eq:pmi}) (see e.g.~\newcite{J90-1003}, \newcite{Turney:2010:FMV:1861751.1861756}, \newcite{NIPS2014_5477}).
%
\begin{equation}
  \label{eq:pmi}
  \operatorname{PMI}(x, y) = \log\frac{P(x,y)}{P(x)P(y)}
\end{equation}

PMI in its raw form is problematic: non-observed co-occurrences lead to infinite PMI values, making it impossible to compute similarity. A common ``fix'' to this problem is to replace all infinities with zeroes, and we use PMI hereafter to refer to a weighting with this fix.

An alternative solution is to increment the probability ratio by 1; we refer to this as \textit{compressed PMI} (CPMI):
%
\begin{equation}
  \label{eq:cpmi}
  \operatorname{CPMI}(x, y) = \log\Big( 1 + \frac{P(x,y)}{P(x)P(y)} \Big)
\end{equation}

@@

\newcite{TACL570} show that performance is affected by smoothing the context distribution $P(x)$:
%
\begin{equation}
  \label{eq:cds}
  P_{\alpha}(x) = \frac{n(x)^{\alpha}}{\sum_{c}n(c)^{\alpha}}
\end{equation}
%
We experiment with $\alpha$ values of 1 (no smoothing) and 0.75. We refer to this estimation method as \emph{local context probability}; we can also estimate a \emph{global context probability} based on global context appearance:
%
\begin{equation}
  \label{eq:cds-nan}
  P(x) = \frac{n(x)}{|C|}
\end{equation}
%
where $|C|$ is the size of the used corpus.

@@

PMI should be used with global context probabilities. \NCPMI/ and \logNCPMI/ should be used with local context probabilities and \CPMI/ should apply context distribution smoothing with $\alpha = 0.75$.

@@

We evaluate these heuristics by comparing the performance they give on SimLex-999 against that obtained using the best possible parameter selections (determined via an exhaustive search for each dimensionality choice); and the best models reported by \cite{TACL570} (see Figure~\ref{fig:simlex-ppmi-best-simlex}).

@@

Heuristics for \logNSCPMI/and \SCPMI/ follow the best selection, but with a wider gap than the SPMI models. In general $n$ weighted models do not perform as well as others.

>>>>

In contrast, much work on \emph{compositional} distributional semantics employs vectors with much fewer dimensions: e.g.~\textbf{2K} \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, \textbf{3K} \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or \textbf{10K} \cite{polajnar-clark:2014:EACL,Baroni2010nouns}. These models  assign tensors to functional words; for a vector space $V$ with dimension $k$,  a tensor $V \otimes V \cdots \otimes V$ of rank $n$ has dimension  $n^k$. As intransitive verbs and adjectives have  tensors of rank 2,  transitive verbs  rank 3, and prepositions rank 7, 
taking $n= \textbf{30K}$ results in spaces with intractable dimensions  $\textbf{30K}^\textbf{2}$, $\textbf{30K}^\textbf{3}$ and $\textbf{30K}^\textbf{7}$.

@@

Most co-occurrence weighting schemes in distributional semantics are based on \emph{point-wise mutual information} (PMI, Equation~\ref{eq:pmi}) (see e.g.~\newcite{J90-1003}, \newcite{Turney:2010:FMV:1861751.1861756}, \newcite{NIPS2014_5477}).
%
\begin{equation}
  \label{eq:pmi}
  \operatorname{PMI}(x, y) = \log\frac{P(x,y)}{P(x)P(y)}
\end{equation}
%
PMI in its raw form is problematic: non-observed co-occurrences lead to infinite PMI values, making it impossible to compute similarity. A common ``fix'' to this problem is to replace all infinities with zeroes, and we use PMI hereafter to refer to a weighting with this fix.
%
An alternative solution is to increment the probability ratio by 1; we refer to this as \textit{compressed PMI} (CPMI):
%
\begin{equation}
  \label{eq:cpmi}
  \operatorname{CPMI}(x, y) = \log\Big( 1 + \frac{P(x,y)}{P(x)P(y)} \Big)
\end{equation}

@@

\newcite{TACL570} show that performance is affected by smoothing the context distribution $P(x)$:
%
\begin{equation}
  \label{eq:cds}
  P_{\alpha}(x) = \frac{n(x)^{\alpha}}{\sum_{c}n(c)^{\alpha}}
\end{equation}
%
We experiment with $\alpha$ values of 1 (no smoothing) and 0.75. We refer to this estimation method as \emph{local context probability}; we can also estimate a \emph{global context probability} based on the size of the corpus $C$:
%
\begin{equation}
  \label{eq:cds-nan}
  P(x) = \frac{n(x)}{|C|}
\end{equation}
%
%where $|C|$ is the size of the used corpus.

@@

PMI should be used with global context probabilities. CPMI generally outperforms PMI, with less sensitivity to parameters; \NCPMI/ and \logNCPMI/ should be used with local context probabilities and \CPMI/ should apply context distribution smoothing with $\alpha = 0.75$.

@@

We evaluate these heuristics by comparing the performance they give on SimLex-999 against that obtained using the best possible parameter selections (determined via an exhaustive search at each dimensionality setting); as well as the best scores reported by \cite{TACL570}, and those given by word2vec-SGNS \cite{mikolov2013efficient} and GloVe \cite{pennington2014glove} -- see Figure~\ref{fig:simlex-ppmi-best-simlex}.

@@

Heuristics for \logNSCPMI/ and \SCPMI/ follow the best selection, but with a wider gap than the SPMI models. In general $n$ weighted models do not perform as well as others.
